---
title: "R Notebook"
output: html_notebook
---

```{r}
library(dplyr)
library(ggplot2)
library(stargazer)
install.packages('truncreg')
library(truncreg) #truncated regression model
install.packages('censReg')
library(censReg) #Tobit model, censored regression model
install.packages('crch')
library(crch) #censored regression with conditional heteroskedasticity, flexible models
load(url("http://kodu.ut.ee/~avork/files/oppetoo/micro/piaacHWTobitHeckmanPII2020.Rdata")) 
```

```{r}
#Analyse distribution of hours worked, fraction of zero hours, distribution of hourly wages, fraction of missing wages
hist(dfr$whours)
hist(dfr$whours[dfr$whours>0])
prop.table(table(dfr$whours==0))
hist(dfr$lwageh)

```

```{r}
library(censReg) 
mytobit<- censReg( whours ~ age + edcat + immig + goodhealth + mathskill1 + children,  left=0, data=dfr )
summary(mytobit)

```
Positive coefficients mean that latent dependent variable y
i increases if x increases
Probability that y > 0 increases and yi increases if x increases
edcat sec is not in ex direction
children is not in the expected direction
```{r}
me.mytobit<- margEff(mytobit) 
summary(me.mytobit)

```
If explanatory variable increases by one unit, how much observed value, conditional that it
is not censored, increases on average
If kids increase increases by one unit, then observed hours incr by 0..86 hours  per week
```{r}
#probit model
dfr$whourspos<-(dfr$whours>0)*1
myprobit<- glm(whourspos ~ age + edcat + immig + goodhealth + mathskill1 + children, family=binomial(link="probit"), data=dfr )
summary(myprobit)

```

```{r}
#Truncated regression
library(truncreg)
mytruncreg<- truncreg(whours ~ age + edcat + immig + goodhealth + mathskill1 + children, point=0, direction= "left", subset =whours>0, data=dfr)
summary(mytruncreg)

```

```{r}
round(cbind(
  "Tobit"     = coef(mytobit)[1:15] /exp(coef(mytobit)[16]),
  "Binary"    = coef(myprobit),
  "Truncated" = coef(mytruncreg)[1:15] / coef(mytruncreg)[16]),4)

```
tobit model is not reasonable
```{r}
LRtest <- as.numeric(2*(logLik(myprobit) +  logLik(mytruncreg) - logLik(mytobit)))
#Difference in degrees of freedom:
dof= length(coefficients(myprobit)) + length(coefficients(mytruncreg)) - length(coefficients(mytobit))
pchisq(LRtest, dof, lower.tail = FALSE)

```
tobit is not fine the two part flexible model is better
```{r}
install.packages('sampleSelection')
library(sampleSelection)
myheckman <- selection(whourspos ~ age + edcat  + immig + mathskill1 + goodhealth + children, lwageh ~ age + edcat + immig + mathskill1 , data=dfr)
summary(myheckman)

```
yes it has the expected signs.
If education increases by one year,
then average wage increases
10.8% (or 10.9% in the two-step
estimator)
those with uppersec are less likely to partcipate and postsec more likely to participate
both earn more compared to lowersec.
wages increases with age
mathskills increases wages
we dont need heckman selection models
```{r}
myolswage <- lm(lwageh ~ age + male + edcat + geo + immig + mathskill1 , data=dfr) 
stargazer(myheckman, myolswage, type="text", no.space = TRUE)

```
no much difference, we should use ols on observed variables is ok
